\section{Fundamental Theorem of Galois Theory}

Let $L/K$ be a finite Galois extension with group $G=\Gal(L/K)$.

\subsection{Statement}

\begin{theorem}[FTGT]
The maps
\[
\Phi:\{\text{subgroups }H\le G\}\to\{\text{intermediate fields }E,\ K\subseteq E\subseteq L\},\quad H\mapsto L^{H},
\]
\[
\Psi:\{E\}\to\{H\},\quad E\mapsto \Gal(L/E),
\]
are inclusion-reversing inverses. Moreover:
\begin{enumerate}
\item $[E:K]=|G:\Gal(L/E)|$ and $[L:E]=|\Gal(L/E)|$.
\item $E/K$ is Galois $\iff$ $\Gal(L/E)\trianglelefteq G$, and then
\[
\Gal(E/K)\cong G/\Gal(L/E).
\]
\end{enumerate}
\end{theorem}

\begin{proof}[Proof sketch]
Artin's theorem gives $K=L^{G}$. If $H\le G$, then $L/L^{H}$ is Galois with group $H$, hence $|\Gal(L/L^H)|=|H|$ and indices match degrees. For any intermediate $E$, clearly $E\subseteq L^{\Gal(L/E)}$ and equality follows by counting degrees. Normality of $E/K$ corresponds to conjugation-stability of $\Gal(L/E)$.
\end{proof}

\subsection{Consequences}

\begin{corollary}[Galois correspondence algebra]
The lattice of subgroups of $G$ is anti-isomorphic to the lattice of intermediate fields of $L/K$. Intersections and joins correspond to composita and intersections of fields.
\end{corollary}

\begin{corollary}[Fixed field of a normal subgroup]
If $N\trianglelefteq G$, then $L^{N}/K$ is Galois with group $G/N$.
\end{corollary}

\subsection{Worked example: $x^3-2$}

Let $E$ be the splitting field of $x^3-2$ over $\Q$. Then $E=\Q(\sqrt[3]{2},\omega)$ and $G\cong S_3$.
\[
\begin{tikzcd}
& E \arrow[-]{dl} \arrow[-]{d} \arrow[-]{dr} & \\
\Q(\sqrt[3]{2}) \arrow[-]{dr} & \Q(\omega) \arrow[-]{d} & \Q(\sqrt[3]{2}\omega) \arrow[-]{dl} \\
& \Q &
\end{tikzcd}
\]
The three index-$2$ subgroups of $S_3$ fix the three quadratic subfields, while the three index-$3$ subgroups fix the three cubic subfields. Complex conjugation corresponds to a transposition.

\subsection{Another example: a quartic with dihedral group}

For $f=x^4-2$ over $\Q$, the splitting field $E=\Q(\sqrt[4]{2},i)$ has $G\cong D_4$ of order $8$. The unique Klein $V_4\trianglelefteq D_4$ fixes $\Q(\sqrt{2})$, hence $\Q(\sqrt{2})/\Q$ is Galois with group $C_2$, and $E/\Q(\sqrt{2})$ is a biquadratic extension.

\subsection{Remarks}

\begin{remark}
The correspondence fails without separability or normality: in general, taking fixed fields of subgroups of \emph{all} automorphisms of $L$ over $K$ still yields intermediate fields, but surjectivity and index-degree equalities can fail.
\end{remark}
