\begin{example}\label{ex:sec9-1}
\textbf{Problem.} Solve $x^2-2=0$ by radicals.

\textbf{Solution.} $x=\pm\sqrt{2}$ solves the equation.
\end{example}

\begin{example}\label{ex:sec9-2}
\textbf{Problem.} Solve $x^3-2=0$ by radicals.

\textbf{Solution.} $x=\sqrt[3]{2}$ and its complex multiples are solutions.
\end{example}

\begin{example}\label{ex:sec9-3}
\textbf{Problem.} Are all quartic equations solvable by radicals?

\textbf{Solution.} Yes, there is a general quartic formula.
\end{example}

\begin{example}\label{ex:sec9-4}
\textbf{Problem.} Are all quintic equations solvable by radicals?

\textbf{Solution.} No, some have Galois group $S_5$ and are not solvable.
\end{example}

\begin{example}\label{ex:sec9-5}
\textbf{Problem.} Is $x^5-1=0$ solvable by radicals?

\textbf{Solution.} Yes, roots are $5$th roots of unity.
\end{example}

\begin{example}\label{ex:sec9-6}
\textbf{Problem.} What does solvable Galois group imply about the polynomial?

\textbf{Solution.} The polynomial is solvable by radicals.
\end{example}

\begin{example}\label{ex:sec9-7}
\textbf{Problem.} Give a polynomial with Galois group $S_3$.

\textbf{Solution.} $x^3-2$ has Galois group $S_3$.
\end{example}

\begin{example}\label{ex:sec9-8}
\textbf{Problem.} Provide a polynomial with Galois group $S_5$.

\textbf{Solution.} $x^5-6x+3$ has Galois group $S_5$ over $\mathbb{Q}$.
\end{example}

\begin{example}\label{ex:sec9-9}
\textbf{Problem.} Does solvability by radicals depend on characteristic $0$?

\textbf{Solution.} Yes, classical results assume characteristic $0$.
\end{example}

\begin{example}\label{ex:sec9-10}
\textbf{Problem.} Give a quintic that is solvable by radicals.

\textbf{Solution.} $x^5-1$ is solvable since its Galois group is cyclic.
\end{example}
