\begin{example}\label{ex:sec7-1}
\textbf{Problem.} How many elements does $\mathbb{F}_5$ have?

\textbf{Solution.} It has $5$ elements.
\end{example}

\begin{example}\label{ex:sec7-2}
\textbf{Problem.} What is the size of any finite field?

\textbf{Solution.} Every finite field has $p^n$ elements for a prime $p$ and integer $n$.
\end{example}

\begin{example}\label{ex:sec7-3}
\textbf{Problem.} Describe the multiplicative group of $\mathbb{F}_5$.

\textbf{Solution.} Nonzero elements form a cyclic group of order $4$.
\end{example}

\begin{example}\label{ex:sec7-4}
\textbf{Problem.} Give a construction of $\mathbb{F}_4$.

\textbf{Solution.} $\mathbb{F}_4 \cong \mathbb{F}_2[x]/(x^2+x+1)$.
\end{example}

\begin{example}\label{ex:sec7-5}
\textbf{Problem.} How many subfields of $\mathbb{F}_{p^n}$ have size $p^m$ dividing $n$?

\textbf{Solution.} Exactly one such subfield exists for each divisor $m$ of $n$.
\end{example}

\begin{example}\label{ex:sec7-6}
\textbf{Problem.} In $\mathbb{F}_4$, compute $(\alpha+1)^2$ for $\alpha^2+\alpha+1=0$.

\textbf{Solution.} $(\alpha+1)^2=\alpha^2+2\alpha+1=\alpha+1$ since $2=0$.
\end{example}

\begin{example}\label{ex:sec7-7}
\textbf{Problem.} What is the Frobenius map on $\mathbb{F}_{p^n}$?

\textbf{Solution.} It is $x\mapsto x^p$, an automorphism.
\end{example}

\begin{example}\label{ex:sec7-8}
\textbf{Problem.} Find the order of $\mathbb{F}_{7^2}^\times$.

\textbf{Solution.} It has $7^2-1=48$ elements.
\end{example}

\begin{example}\label{ex:sec7-9}
\textbf{Problem.} Does $x^{p^n}-x$ split over $\mathbb{F}_{p^n}$?

\textbf{Solution.} Yes, every element is a root, so it splits completely.
\end{example}

\begin{example}\label{ex:sec7-10}
\textbf{Problem.} What is the minimal polynomial of $\alpha$ satisfying $\alpha^2+\alpha+1=0$ over $\mathbb{F}_2$?

\textbf{Solution.} It is $x^2+x+1$, irreducible over $\mathbb{F}_2$.
\end{example}
