\begin{example}\label{ex:sec14-1}
\textbf{Problem.} Does the diagram $A\xrightarrow{\mathrm{id}}A$ commute?

\textbf{Solution.} Yes, the single identity map trivially commutes.
\end{example}

\begin{example}\label{ex:sec14-2}
\textbf{Problem.} In a triangle $A\to B\to C$ with compositions, does $(A\to C)$ equal the composite?

\textbf{Solution.} Yes, by definition of composition.
\end{example}

\begin{example}\label{ex:sec14-3}
\textbf{Problem.} What matrix represents $90^\circ$ rotation in the plane?

\textbf{Solution.} $\begin{pmatrix}0&-1\\1&0\end{pmatrix}$.
\end{example}

\begin{example}\label{ex:sec14-4}
\textbf{Problem.} What is the adjacency matrix of a single edge connecting two vertices?

\textbf{Solution.} $\begin{pmatrix}0&1\\1&0\end{pmatrix}$.
\end{example}

\begin{example}\label{ex:sec14-5}
\textbf{Problem.} Does a square of zero maps commute?

\textbf{Solution.} Yes, all compositions are zero.
\end{example}

\begin{example}\label{ex:sec14-6}
\textbf{Problem.} Draw a short exact sequence diagram.

\textbf{Solution.} $0\to A\to B\to C\to0$ with exact arrows.
\end{example}

\begin{example}\label{ex:sec14-7}
\textbf{Problem.} Can inclusion maps $K\subset L\subset M$ form a commutative triangle?

\textbf{Solution.} Yes, inclusions compose to the larger inclusion.
\end{example}

\begin{example}\label{ex:sec14-8}
\textbf{Problem.} Does the field diagram $\mathbb{Q}\subset\mathbb{Q}(\sqrt{2})\subset\mathbb{Q}(\sqrt{2},i)$ commute?

\textbf{Solution.} Yes, inclusions compose correctly.
\end{example}

\begin{example}\label{ex:sec14-9}
\textbf{Problem.} For group homomorphisms $G\xrightarrow{f}H\xrightarrow{g}K$, does $g\circ f$ define a commuting triangle with $G\to K$?

\textbf{Solution.} Yes, the composite map commutes.
\end{example}

\begin{example}\label{ex:sec14-10}
\textbf{Problem.} Does a commutative square of abelian groups ensure equal composites?

\textbf{Solution.} Yes, commutativity means both paths give the same map.
\end{example}
