\begin{example}\label{ex:sec12-1}
\textbf{Problem.} Can a segment of length $\sqrt{2}$ be constructed?

\textbf{Solution.} Yes, it is the hypotenuse of a unit right triangle.
\end{example}

\begin{example}\label{ex:sec12-2}
\textbf{Problem.} Is $\sqrt[3]{2}$ constructible by ruler and compass?

\textbf{Solution.} No, it has degree $3$, not a power of $2$.
\end{example}

\begin{example}\label{ex:sec12-3}
\textbf{Problem.} Is a regular pentagon constructible?

\textbf{Solution.} Yes, since $5$ is a Fermat prime.
\end{example}

\begin{example}\label{ex:sec12-4}
\textbf{Problem.} Is a regular heptagon constructible?

\textbf{Solution.} No, $7$ is not a Fermat prime.
\end{example}

\begin{example}\label{ex:sec12-5}
\textbf{Problem.} Can a square be doubled exactly?

\textbf{Solution.} Yes, by constructing $\sqrt{2}$.
\end{example}

\begin{example}\label{ex:sec12-6}
\textbf{Problem.} Can a general angle be trisected with ruler and compass?

\textbf{Solution.} No, this is impossible in general.
\end{example}

\begin{example}\label{ex:sec12-7}
\textbf{Problem.} Can an angle be bisected?

\textbf{Solution.} Yes, classical constructions bisect any angle.
\end{example}

\begin{example}\label{ex:sec12-8}
\textbf{Problem.} Can a perpendicular be erected from a point on a line?

\textbf{Solution.} Yes, using circles of equal radius.
\end{example}

\begin{example}\label{ex:sec12-9}
\textbf{Problem.} Is a regular $17$-gon constructible?

\textbf{Solution.} Yes, $17$ is a Fermat prime.
\end{example}

\begin{example}\label{ex:sec12-10}
\textbf{Problem.} Can one construct the angle $\pi/3$?

\textbf{Solution.} Yes, equilateral triangle gives $60^{\circ}$.
\end{example}
