\section{Study Plan and Roadmap}

\subsection*{Prerequisites}
\begin{itemize}
  \item Linear algebra: eigenvalues, minimal polynomials, Jordan form.
  \item Basic ring and field theory: ideals, PID/UFD, polynomial rings.
  \item Group theory: actions, Sylow, simple groups $A_n,S_n$.
\end{itemize}

\subsection*{Learning Outcomes}
\begin{itemize}
  \item Compute splitting fields and Galois groups for key polynomials.
  \item Apply the Fundamental Theorem of Galois Theory (FTGT).
  \item Decide solvability by radicals and explain Abel--Ruffini.
  \item Work with finite fields and infinite Galois groups (Krull topology).
\end{itemize}

\subsection*{Graduate-Level Outline}
\begin{enumerate}
  \item Fields and morphisms; irreducibility tests.
  \item Extensions; algebraic vs.\ transcendental; minimal polynomials.
  \item Splitting fields and algebraic closures.
  \item Separable/inseparable; perfect fields.
  \item Normal extensions; Primitive Element Theorem.
  \item Field automorphisms; Galois extensions and fixed fields.
  \item FTGT: lattice correspondence; normality and separability conditions.
  \item Classical examples: quadratics, cyclotomics, Kummer, finite fields.
  \item Solvability by radicals; solvable groups; Abel--Ruffini.
  \item Discriminants and resolvents.
  \item Finite fields $\F_{p^n}$; Frobenius and Galois group structure.
  \item Infinite Galois theory; profinite groups; Krull topology.
  \item Cyclotomic fields; basic results and computations.
  \item Directions for research and advanced topics (brief notes).
\end{enumerate}

\subsection*{Study Rhythm}
\begin{itemize}
  \item Problem-first: compute Galois groups for degrees $2$--$5$.
  \item Diagram-first: use \texttt{tikz-cd} to track subfield lattices.
  \item Proof drills: restate and prove FTGT variants.
\end{itemize}
